Collapse of tall granular columns in fluid

### 2016-11-21 paper.tex
Abstract and introduction

\author{\firstname{Krishna} \lastname{Kumar}\inst{1}\fnsep\thanks{\email{kks32@cam.ac.uk}} \and
        \firstname{Kenichi} \lastname{Soga}\inst{2}\fnsep\thanks{\email{soga@berkeley.edu}} \and
        \firstname{Jean-Yves} \lastname{Delenne}\inst{3}\fnsep\thanks{\email{}} \and
        \firstname{Farhang} \lastname{Radjai}\inst{4}\fnsep\thanks{\email{}}
        % etc.
}

\institute{Department of Engineering, University of Cambridge, UK 
\and
           University of California, Berkeley, US
\and
           INRA, France
\and 
           CNRS, France
          }

\abstract{%
  Avalanches, landslides, and debris flows are geophysical hazards, which involve rapid mass movement
  of granular solids, water, and air as a single phase system. In order to describe the mechanism of
  immersed granular flows, it is important to consider both the dynamics of the solid phase and the
  role of the ambient fluid~\cite{Denlinger2001}. The dynamics of the solid phase alone are
  insufficient to describe the mechanism of granular flows in fluid. It is important to consider the
  effect of hydrodynamic forces that reduce the weight of the solids inducing a transition from
  dense-compacted to dense-suspended flows, and the drag interactions which counteract the movement
  of the solids~\cite{Meruane2010}. Transient regimes characterised by a change in the solid
  fraction, dilation at the onset of flow and the development of excess pore-pressure, result in
  altering the balance between the stress carried by the fluid and that carried by the grains, thereby
  changing the overall behaviour of the flow. In the present study, a two-dimensional coupled Lattice
  Boltzmann LBM – DEM technique is developed to understand the micro-scale rheology of granular
  flows in fluid.

  In the present study, the collapse of a granular column in fluid is studied using 2D LBM - DEM.
  The effect of initial aspect ratio on the run-out behaviour is investigated. The flow kinematics are
  compared with the dry and buoyant granular collapse to understand the influence of hydrodynamic
  forces and lubrication on the run-out. In the case of tall columns, the amount of material
  destabilised above the failure plane is larger than that of short columns. Hence in tall columns,
  the surface area of the mobilised mass that interacts with the surrounding fluid is significantly
  higher than the short columns. This increase in the area of soil - fluid interaction results in an
  increase in the formation of turbulent vortices that alter the deposit morphology during the
  collapse. It is observed that the vortices result in formation of heaps that significantly affect
  the distribution of mass in the flow.~\cite{Staron2007a} observed that the distribution of mass in
  a granular flow plays a crucial role in the flow kinematics. In order to understand the behaviour
  of tall columns, the run-out behaviour of a dense granular column with an initial aspect ratio of
  6 is studied. The collapse of a tall granular column on slopes of 0, 2.5, 5 and 7.5 are studied.
  The permeability of the granular mass was varied to understand the effect of permeability on the
  run-out behaviour. }
%
\maketitle
%
\section{Introduction}
The collapse of a granular column on a horizontal surface is a simple case
of granular flow, however a proper model that describes the flow dynamics
is still lacking. Experimental investigations have shown that the flow
duration, the spreading velocity, the final extent of the deposit, and the
energy dissipation can be scaled in a quantitative way independent of
substrate properties, grain size, density, and shape of the granular
material and released mass~\cite{Lajeunesse2005, Lube2005}. Simple
mathematical models based on conservation of horizontal momentum capture
the scaling laws of the final deposit, but fail to describe the initial
transition regime. Granular flow is modelled as a frictional dissipation
process in continuum mechanics but the lack of influence of inter-particle
friction on the energy dissipation and spreading dynamics~\cite{Lube2005}
is surprising.

>>>>

\author{\firstname{Krishna} \lastname{Kumar}\inst{1}\fnsep\thanks{\email{kks32@cam.ac.uk}} \and
        \firstname{Kenichi} \lastname{Soga}\inst{2}\fnsep\thanks{\email{soga@berkeley.edu}} \and
        \firstname{Jean-Yves} \lastname{Delenne}\inst{3}\fnsep\thanks{\email{}}
}

\institute{Department of Engineering, University of Cambridge, UK 
\and
           University of California, Berkeley, USA
\and
           INRA, France
}

\abstract{%
  Avalanches, landslides, and debris flows are geophysical hazards, involve
  rapid mass movement of granular solids, water, and air as a multi-phase
  system. In order to describe the mechanism of immersed granular flows,
  it is important to consider both the dynamics of the solid phase and the
  role of the ambient fluid~\cite{Denlinger2001}. In the present study,
  the collapse of a granular column in fluid is
  studied using 2D LBM - DEM. The flow kinematics are compared with the dry
  and buoyant granular collapse to understand the influence of hydrodynamic
  forces and lubrication on the run-out. In the case of tall columns,
  the amount of material destabilised above the failure plane is larger than
  that of short columns. Hence in tall columns, the surface area of the
  mobilised mass that interacts with the surrounding fluid is significantly
  higher than the short columns. This increase in the area of soil - fluid
  interaction results in an increase in the formation of turbulent vortices
  that alter the deposit morphology during the collapse. It is observed that
  the vortices result in the formation of heaps that significantly affect
  the distribution of mass in the flow. In order to understand the behaviour
  of tall columns, the run-out behaviour
  of a dense granular column with an initial aspect ratio of 6 is studied.
  The collapse of a tall granular column on slopes of 0, 2.5, 5 and 7.5
  are studied.}
%
\maketitle
%
\section{Collapse of granular columns}
The collapse of a granular column on a horizontal surface is a simple case
of granular flow, however a proper model that describes the flow dynamics
is still lacking. Experimental investigations have shown that the flow
duration, the spreading velocity, the final extent of the deposit, and the
energy dissipation can be scaled in a quantitative way independent of
substrate properties, grain size, density, and shape of the granular
material and released mass~\cite{Staron2007a, Lajeunesse2005, Lube2005}.

Granular collapse on a horizontal plane exhibit two distinct flow regimes:
(a) for columns with aspect ratio ‘a’ <1.7 a linear relation between the
spread and aspect ratio can be observed, and (b) for ‘a’ > 1.7 a power-law
relationship exists.~\citet{Soundararajan2015} observed that a simple
frictional dissipation model is able to capture the flow dynamics for
columns with small aspect ratios. However, for tall columns the flow
dyanmics of is controlled by the initial collisional regime, resulting in a
power-law relationship with the aspect ratios. Simple mathematical models
based on conservation of horizontal momentum capture the scaling laws of
the final deposit, but fail to describe the initial transition regime
~\citep{Kumar2013}. Granular flow is modelled as a frictional dissipation
process in continuum mechanics but the lack of influence of inter-particle
collision and friction on the energy dissipation and spreading dynamics is
surprising.
